Simple and robust inference of connectomes using partial correlation coefficients

### 2014-06-09 article.tex
fix causal notation
break at 80

Then we apply a non-linear filter based on the current activity of the systems
in order to favor the association measure when a medium size set of neurons
are firing together, while artificially reducing the association measure when
the activity is extremely high, i.e. in case of a network burst.

\begin{align}
h_1(X_t^i) = (X_t^i + 1)^{\left( 1 + \dfrac{1}{a_t} \right)} &\text{ with } a_t = \sum_{i=1}^p \left ( X_t^i + \dfrac{1}{2} X_{t-1}^i \right ) \\
h_2(X_t^i) = (X_t^i)^{r(a_t)} &\text{ with } r_t = \frac{a_t}{\max_{l \in T}{a_l}} \\
\end{align}

@@

% Filtering methodology
Prior inferring the network, the fluorescent calcium time series are smoothed
using several linear and non linear filters. A low-pass filter, a high-pass
filter and a non linear hard-thresholding based filter are
applied for each possible time step $t$
% TODO fix first equation is wrong due notation change
\begin{align}
X^\prime_{t,i} &= \sum_{l=-3}^2 c_l^k X_{t+l,i} \forall i, \\
X^{\prime\prime}_{t,i} &= X^{\prime}_{t,i} - X^{\prime}_{t-1,i} \forall i, \\
X^{\prime\prime\prime}_{t,i} &= {X^{\prime\prime}_{t,i}}^{0.9} I(X^{\prime\prime}_{t,i} \geq f_3) \forall i,
\end{align}
where the $c_l^k$ coefficients were taken in one of the four following tuples
$\left\{c_l^1\right\}=(0, 0, 1, 1, 1, 1)$,
$\left\{c_l^2\right\}=(0, 0, 1, 1, 1, 0)$,
$\left\{c_l^3\right\}=(1, 1, 1, 1, 0, 0)$ or
$\left\{c_l^4\right\}=(0.4, 0.8, 1, 1, 0, 0)$ and $f_3$ is a parameter
of the hard-threshold filter.

@@

\textcolor{red}{Motivation pour l'inverse : retirer les liens indirects. Pareil: proposition de AS  et certains choses pourraient être incorporées dans la version de AJ.} \\

@@

Each neuron signal $i$ is a random continuous variable $X_i$,
which follows a probability density function $p_\mathcal{X_i}(x_i)$.
Two neurons $X_i$, $X_j$ will be said conditionally independent
to a set of neurons $\mathcal{V}^{-i,j}$, denoted
by  $X_i \perp X_j | \mathcal{V}^{-i,j}$,  if $p_{X_i,X_j|\mathcal{V}^{-i,j}} = p_{X_i|\mathcal{V}^{-i,j}}  p_{X_j|\mathcal{V}^{-i,j}} $.
%for all real numbers $x_1,\ldots,x_p $ such that
%$p_{X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}}
%(x_1,\ldots,x_p \backslash \left\{x_i, x_j\right\}) > 0$. Note that this
%reciprocal is true only under additional hypotheses.

@@

By ignoring the directivity, we systematically
make a true positive and a false positive for direct link (see a more detailed
discussion about the chosen metric in section \ref{sec:metric}). Therefore,
performance might be improved by predicting an asymmetric adjacency matrix.
Hoping to orientate correctly the most obvious directed links, we attempt to
retrieve some causal information, i.e. the directivity of the links, by computing
and stacking with our symmetric adjacency matrix a matrix of pairwise activation
heuristics. This method computes an anti-symmetric activation score $z_{ij}$
based the variation of fluorescence signal of a neuron $j$ due to a neuron $i$
over all timesteps:
\[
z_{ij} = \sum_{t=1}^n 1(X_{t+1,j} \ge X_{t, i} \in [f_1, f_2]) -  1(X_{t+1,i} \ge X_{t, j} \in [f_1, f_2])
\] \textcolor{red}{to rewrite}
where $f_1$ and $f_2$ are parameters of the method and $I$ is the
indicator function. % TODO say that we use only the 3 first filters for this

>>>>

Then we apply a non-linear filter based on the current activity of the systems
in order to favor the association measure when a medium size set of neurons
are firing together, while artificially reducing the association measure when
the activity is extremely high, i.e. in case of a network burst. \begin{align}
h_1(X_t^i) = (X_t^i + 1)^{\left( 1 + \dfrac{1}{a_t} \right)} &\text{ with }
a_t = \sum_{i=1}^p \left ( X_t^i + \dfrac{1}{2} X_{t-1}^i \right ) \\
h_2(X_t^i) = (X_t^i)^{r(a_t)} &\text{ with } r_t = \frac{a_t}{\max_{l \in
T}{a_l}}  \end{align}

@@

% Filtering methodology Prior inferring the network, the fluorescent calcium
time series are smoothed using several linear and non linear filters. A low-
pass filter, a high-pass filter and a non linear hard-thresholding based
filter are applied for each possible time step $t$ % TODO fix first equation
is wrong due notation change \begin{align} X^\prime_{t,i} &= \sum_{l=-3}^2
c_l^k X_{t+l,i} \forall i, \\ X^{\prime\prime}_{t,i} &= X^{\prime}_{t,i} -
X^{\prime}_{t-1,i} \forall i, \\ X^{\prime\prime\prime}_{t,i} &=
{X^{\prime\prime}_{t,i}}^{0.9} I(X^{\prime\prime}_{t,i} \geq f_3) \forall i,
\end{align} where the $c_l^k$ coefficients were taken in one of the four
following tuples $\left\{c_l^1\right\}=(0, 0, 1, 1, 1, 1)$,
$\left\{c_l^2\right\}=(0, 0, 1, 1, 1, 0)$, $\left\{c_l^3\right\}=(1, 1, 1, 1,
0, 0)$ or $\left\{c_l^4\right\}=(0.4, 0.8, 1, 1, 0, 0)$ and $f_3$ is a
parameter of the hard-threshold filter.

@@

\textcolor{red}{Motivation pour l'inverse : retirer les liens indirects.
\Pareil: proposition de AS  et certains choses pourraient être incorporées
\dans la version de AJ.}

@@

Each neuron signal $i$ is a random continuous variable $X_i$, which follows a
probability density function $p_\mathcal{X_i}(x_i)$. Two neurons $X_i$, $X_j$
will be said conditionally independent to a set of neurons
$\mathcal{V}^{-i,j}$, denoted by  $X_i \perp X_j | \mathcal{V}^{-i,j}$,  if
$p_{X_i,X_j|\mathcal{V}^{-i,j}} = p_{X_i|\mathcal{V}^{-i,j}}
p_{X_j|\mathcal{V}^{-i,j}} $.

%for all real numbers $x_1,\ldots,x_p $ such that
%$p_{X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}}
%(x_1,\ldots,x_p \backslash \left\{x_i, x_j\right\}) > 0$. Note that this
%reciprocal is true only under additional hypotheses.

@@

By ignoring the directivity, we systematically make a true positive and a
false positive for direct link (see a more detailed discussion about the
chosen metric in section \ref{sec:metric}). Therefore, performance might be
improved by predicting an asymmetric adjacency matrix. Hoping to orientate
correctly the most obvious directed links, we attempt to retrieve some causal
information, i.e. the directivity of the links, by computing and stacking with
our symmetric adjacency matrix a matrix of pairwise activation heuristics.
This method computes an anti-symmetric activation score $z_{ij}$ based the
variation of fluorescence signal of a neuron $j$ due to a neuron $i$ over all
timesteps: \[ z_{ij} = \sum_{t=1}^n \mathbb{1}(X_{t+1,j} \ge X_{t, i}) \cdot
\mathbb{1}(f_1\le X_t^i \le f_2) -  \mathbb{1}(X_{t+1,i} \ge X_{t, j}) \cdot
\mathbb{1}(f_1 \le X_t^j \le f_2) \]  where $f_1$ and $f_2$ are parameters of
the method and $I$ is the indicator function. % TODO say that we use only the
3 first filters for this
